---
title: "Agriculture Dataset Exploratory Analysis"
description: |
  Exploratory analysis of the crop production data source. 
author:
  - name: Aditi Gautam
    url: https://github.com/adgautam
date: "`r Sys.Date()`"
output: 
    radix::radix_article:
      toc: true
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	rows.print=5
)

options(tibble.max_extra_cols = 5, tibble.print_max = 5)

library(fs) # cross-platform file system operations
library(tidyverse)
library(skimr)
```

#Importing dataset

```{r}
agri <- read_csv(path("apy.csv"))
```

#Exploring and summarizing the dataset

```{r layout="l-body-outset"}
agri
```

```{r}
#library(dplyr)
glimpse(agri)
```

```{r}
#library(skimr)
skim(agri)
```

#Creating column for calculating Production tonnes per hectare and to identify MSP crops 

```{r}
agri <- agri %>% mutate(
  Production_by_Area = Production / Area,
  msp = Crop %in% c("Paddy", "Wheat", "Maize", "Barley", "Ragi", "Gram", "Arhar/Tur", "Moong(Green Gram)", "Urad", "Masoor", "Groundnut", "Rapeseed &Mustard", "Soyabean", "Sesamum", "Sunflower", "Niger seed", "Cotton", "Jute & mesta", "Jowar", "Bajra", "Sugarcane")) 

agri
```

#Creating Average production for MSP and Non-MSP crops by year

```{r}
agri_sum <- agri %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(mean = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_sum
```

#Graphing average production tonnes per hectare

```{r}
ggplot(agri_sum) + 
  aes(x = Crop_Year, y = mean, color = msp) + 
  geom_line() + 
  theme_bw()
```
#Looking at outliers in MSP vs non-MSP crops. Many non-MSP crops seem to be outliers.

```{r}
ggplot(filter(agri, !is.na(Production_by_Area))) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = msp) +
  geom_boxplot() +
  theme_bw()
```


```{r}
group_by(agri, Crop_Year, Crop) %>%
  summarise(mean = scales::comma(mean(Production, na.rm = TRUE), accuracy = 1)) %>%
  pivot_wider(
    names_from = Crop_Year,
    values_from = mean
  )
```

```{r}
agri %>% filter(msp == TRUE, log(Production_by_Area) > 10) #there seems to be an issue with the data for Area in Punjab in 2011 leading to the big jump seen in the boxplot for MSP crops. 
```
#I will now be analysing data by 1. removing Punjab for Sugarcane for 2011. 2. Multiplying Punjab area by 1000 to correct what we assume is an error in the data. 

#1. Removing Punjab-Sugarcane for 2011
```{r}
agri_nopunj <- agri[agri$Crop_Year != 2011 & agri$Crop != "Sugarcane", ]  
ggplot(filter(agri_nopunj, !is.na(Production_by_Area))) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = msp) +
  geom_boxplot() +
  theme_bw()
```
```{r}
agri_nopunj_sum <- agri_nopunj %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(mean = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_nopunj_sum
```


```{r}
ggplot(agri_nopunj_sum) + 
  aes(x = Crop_Year, y = mean, color = msp) + 
  geom_line() + 
  theme_bw()
```


#Looking at outliers in production for non-msp crops

```{r}
agri_nopunj %>% 
  group_by(Crop, Crop_Year) %>%
  filter(msp == FALSE, log(Production_by_Area) > 5) %>%
  ungroup() #Coconut seems to be an outlier that is driving the massive difference between msp and non-msp crops. 
```
```{r}
agri_nopunj %>% 
  group_by(Crop, Crop_Year) %>%
  filter(msp == FALSE, log(Production_by_Area) < -5) %>%
  ungroup()
```


```{r}
agri_nopunj_nococo <- agri_nopunj %>% 
  group_by(Crop, Crop_Year) %>%
  filter(log(Production_by_Area) < 10 & log(Production_by_Area) > -10) %>%
  ungroup()

agri_nopunj_nococo
```
```{r}
agri_nopunj_nococo_sum <- agri_nopunj_nococo %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(mean = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_nopunj_nococo_sum
```


```{r}
ggplot(agri_nopunj_nococo_sum) + 
  aes(x = Crop_Year, y = mean, color = msp) + 
  geom_line() + 
  theme_bw()
```



