---
title: "Agriculture Dataset Exploratory Analysis"
description: |
  Exploratory analysis of the crop production data source. 
author:
  - name: Aditi Gautam
    url: https://github.com/adgautam
date: "`r Sys.Date()`"
output: 
    radix::radix_article:
      toc: true
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	rows.print=5
)

options(tibble.max_extra_cols = 5, tibble.print_max = 5)

library(fs) # cross-platform file system operations
library(tidyverse)
library(skimr)
library(plotly)
```

# Importing dataset

```{r}
agri <- read_csv(path("Agriculture_1997_2015.csv"))
```

# Exploring and summarizing the dataset

```{r layout="l-body-outset"}
agri
glimpse(agri)
skim(agri)
```

# Creating column for calculating Production tonnes per hectare and to identify MSP crops 

```{r}
agri <- agri %>% mutate(
  Production_by_Area = Production / Area,
  msp = Crop %in% c("Paddy", "Wheat", "Maize", "Barley", "Ragi", "Gram", "Arhar/Tur", "Moong(Green Gram)", "Urad", "Masoor", "Groundnut", "Rapeseed &Mustard", "Soyabean", "Sesamum", "Sunflower", "Niger seed", "Cotton", "Jute & mesta", "Jowar", "Bajra", "Sugarcane", "Coconut")) 

agri
```
# Looking and dropping outlier years 

```{r}
aggregate(agri$Production_by_Area, list(agri$Crop_Year), FUN=mean, na.rm=TRUE) # 1997 and 2015 seem like outliers so we drop them. We will see what is happening in 2011 more closely below. 
agri <- agri[agri$Crop_Year != 1997,]
agri <- agri[agri$Crop_Year != 2015,]
aggregate(agri$Production_by_Area, list(agri$Crop_Year), FUN=mean, na.rm=TRUE) 
```

# Creating Average production for MSP and Non-MSP crops by year

```{r}
agri_sum <- agri %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(Average_Production_by_Area = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_sum
```

# Graphing average production tonnes per hectare

```{r}
ggplot(agri_sum) + 
  aes(x = Crop_Year, y = Average_Production_by_Area, color = msp) + 
  geom_line() + 
  labs(
    title = 'Production by Area by Year - MSP vs Non-MSP crops',
    subtitle = "MSP crops have a much greater Production by Area compared to Non-MSP",
    x = "Year",
    y = "Production by Area (tonnes per hectare)",
    caption = "Source: Ministry of Agriculture, India"
  ) +
  theme_bw()
ggsave("Visualizations/Yield1.png")
```

# Looking at outliers in MSP vs non-MSP crops. Many MSP crops seem to be outliers.

```{r}
ggplot(filter(agri, !is.na(Production_by_Area))) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = msp) +
  geom_boxplot() +
  theme_bw() +
   labs(
    title = 'Log of Production by Area by Year - MSP vs Non-MSP crops (Checking for outliers)',
    subtitle = "MSP crops have a much higher Production",
    x = "Year",
    y = "Log of Production by Area (tonnes per hectare)",
    caption = "Source: Ministry of Agriculture, India"
  )
ggsave("Visualizations/Log_Yield1.png")
```

# Looking at production averages by crop

```{r}
group_by(agri, Crop_Year, Crop) %>%
  summarise(mean = scales::comma(mean(Production, na.rm = TRUE), accuracy = 1)) %>%
  pivot_wider(
    names_from = Crop_Year,
    values_from = mean
  )
```

# Looking at outliers in production for msp crops

```{r}
agri %>% 
  group_by(Crop, Crop_Year) %>%
  filter(msp == TRUE, log(Production_by_Area) > 5) %>%
  ungroup()
```

# I will now be analysing data by 1. Removing Punjab for Sugarcane for 2011 along with other outliers. 2. Multiplying Punjab area for sugarcane in 2011 by 1000 to correct what we assume is an error in the data and dropping other outliers. 

# 1. Dropping Punjab-Sugarcane for 2011 and other outliers

```{r}
agri_nooutlier <- agri %>% 
  group_by(Crop, Crop_Year) %>%
  filter(log(Production_by_Area) < 5) %>%
  ungroup()

agri_nooutlier
```

```{r}
ggplot(filter(agri_nooutlier, !is.na(Production_by_Area))) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = msp) +
  geom_boxplot() +
  theme_bw()
ggsave("Visualizations/Log_Yield2.png")
```

# Aggregating data by MSP and Year without Outliers

```{r}
agri_nooutlier_sum <- agri_nooutlier %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(mean = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_nooutlier_sum
```

```{r}
ggplot(agri_nooutlier_sum) + 
  aes(x = Crop_Year, y = mean, color = msp) + 
  geom_line() + 
  theme_bw() +
  labs(
    title = 'Average Production by Area by Year - Without the MSP Outliers',
    subtitle = "MSP and non-MSP have a volatile relationship",
    x = "Year",
    y = "Production by Area (tonnes per hectare)",
    caption = "Source: Ministry of Agriculture, India"
  )
ggsave("Visualizations/Yield2.png")
```

# Renaming state and district columns

```{r}
agri_nooutlier <- agri_nooutlier %>% 
  rename(
    state = State_Name,
    district = District_Name
    )
```


```{r}
agri_nooutlier <- agri_nooutlier %>% mutate(state = toupper(state))
```

# Exporting the cleaned dataset
```{r}
write.csv(agri_nooutlier, file = "Agridata_Cleaned_nopunjab.csv")
```

# 2. Multiplying Punjab area for sugarcane in 2011 by 1000 to correct what we assume is an error in the data and dropping ther outliers

# Correcting Area variable 

```{r}
agri_corr <- agri_nococo

agri_corr <- agri_corr %>%
  mutate( Area = 
    case_when(
      Crop == "Sugarcane" & Crop_Year == 2011 & State_Name == "Punjab" ~  Area*1000, TRUE ~ Area )
  )
```

# Correcting Yield variable

```{r}
agri_corr <- agri_corr %>%
  mutate( Production_by_Area = 
    case_when(
      Crop == "Sugarcane" & Crop_Year == 2011 & State_Name == "Punjab" ~  Production_by_Area/1000, TRUE ~ Production_by_Area )
  )
```

# Dropping rest of outliers - mostly coconut

```{r}
agri_corr <- agri_corr %>% 
  group_by(Crop, Crop_Year) %>%
  filter(log(Production_by_Area) < 5) %>%
  ungroup()

agri_corr
```

```{r}
ggplot(filter(agri_corr, !is.na(Production_by_Area))) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = msp)+
  geom_boxplot() +
  theme_bw()
ggsave("Visualizations/Log_Yield3.png")
```

# Looking at log of production for all MSP crops

```{r}
a<-ggplot(filter(agri_corr, !is.na(Production_by_Area), msp == TRUE)) +
  aes(x = as.character(Crop_Year), y = log(Production_by_Area), color = Crop) +
  geom_boxplot() +
  theme_bw() +   
  xlab("Year") +
  ylab("Crop Yield") +
  labs(
    title = "MSP Crop Yield Across Year",
    caption = "Source: Ministry of Agriculture, India",
  )
a + theme(legend.position="bottom", legend.text = element_text(size=7, face="bold"), legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid"))
ggsave("Visualizations/Log_Yield_Final.png")
```

# Aggregating crop production by MSP and Year

```{r}
agri_corr_sum <- agri_corr %>% 
  group_by(msp, Crop_Year) %>% 
  summarise(mean = mean(Production_by_Area, na.rm=TRUE),
            count = n()) %>% 
  ungroup()

agri_corr_sum
```

# Plotting the Yield by year and MSP

```{r}
ggplot(agri_corr_sum) + 
  aes(x = Crop_Year, y = mean, color = msp) + 
  geom_line() + 
  theme_bw() +
  labs(
    title = 'Average Production by Area by Year - With MSP outlier corrected',
    subtitle = "MSP vs Non-MSP relationship more volatile",
    x = "Year",
    y = "Production by Area (tonnes per hectare)",
    caption = "Source: Ministry of Agriculture, India"
  )
ggsave("Visualizations/Yield_Final.png")
```

# Renaming state and district columns

```{r}
agri_corr <- agri_corr %>% 
  rename(
    state = State_Name,
    district = District_Name
    )
```

```{r}
agri_corr <- agri_corr %>% mutate(state = toupper(state))
```

# Exporting the cleaned dataset
```{r}
write.csv(agri_corr, file = "Agridata_Cleaned_Corrected.csv")
```
